### 2013-07-27 PAPER/ms_final_bv.tex
Added Introduction and plots

% The command below calls the preprint style^M
% which will produce a one-column, single-spaced document.^M
% Examples of commands for other sub-styles follow. Use^M
% whichever is most appropriate for your purposes.^M

%\documentclass[12pt,preprint]{aastex}
%\documentclass[iop,appendixfloats]{emulateapj}

% manuscript produces a one-column, double-spaced document:

%\documentclass[manuscript]{aastex}
%----------------------------------
% mn2esample.tex
%
% v2.1 released 22nd May 2002 (G. Hutton)
%
% The mnsample.tex file has been amended to highlight
% the proper use of LaTeX2e code with the class file
% and using natbib cross-referencing. These changes
% do not reflect the original paper by A. V. Raveendran.
%
% Previous versions of this sample document were
% compatible with the LaTeX 2.09 style file mn.sty
% v1.2 released 5th September 1994 (M. Reed)
% v1.1 released 18th July 1994
% v1.0 released 28th January 1994

\documentclass[useAMS,usenatbib,letters]{mn2e}

% If your system does not have the AMS fonts version 2.0 installed, then
% remove the useAMS option.
%
% useAMS allows you to obtain upright Greek characters.
% e.g. \umu, \upi etc.  See the section on "Upright Greek characters" in
% this guide for further information.
%
% If you are using AMS 2.0 fonts, bold math letters/symbols are available
% at a larger range of sizes for NFSS release 1 and 2 (using \boldmath or
% preferably \bmath).
%
% The usenatbib command allows the use of Patrick Daly's natbib.sty for
% cross-referencing.
%
% If you wish to typeset the paper in Times font (if you do not have the
% PostScript Type 1 Computer Modern fonts you will need to do this to get
% smoother fonts in a PDF file) then uncomment the next line
% \usepackage{Times}

%%%%% AUTHORS - PLACE YOUR OWN MACROS HERE %%%%%
\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\definecolor{orange}{rgb}{1,0.5,0}
\usepackage[colorlinks=True,citecolor=orange,urlcolor=blue,linkcolor=red]{hyperref}

\newcommand{\figurepath}{.}

%  Definition of fonts

\newcommand{\tom}{\tilde{\omega}}

\newcommand{\ri}{r_{\rm i}}

\newcommand{\ro}{r_{\rm out}}

\newcommand{\rs}{r_{\star}}

\newcommand{\mj}{\dot{M}_{\rm jet}}

\newcommand{\ms}{\dot{M}_{\star}}

\newcommand{\vk}{v_{\rm K}}

\newcommand{\ti}{t_{\rm i}}

\newcommand{\rj}{R_{\rm jet}}

\newcommand{\rk}{R_{\kappa}}

\newcommand{\cs}{c_{\rm s}}

\newcommand{\css}{c_{\rm s}^2}

\newcommand{\bp}{B_{\rm P}}

\newcommand{\bh}{B_{\phi}}

\newcommand{\jp}{\vec{B}_{\rm P}}

\newcommand{\cm}{{\rm cm}}

\newcommand{\msun}{{\rm M}_{\sun}}

\newcommand{\rsun}{{\rm R}_{\sun}}

\newcommand{\rstar}{{\rm R}_{\star}}

\newcommand{\lumsun}{{\rm L}_{\sun}}

\newcommand{\lumstar}{{\rm L}_{\star}}

\newcommand{\lumdisk}{{\rm L}_{\rm disk}}

\newcommand{\etat}{{\eta}_{\rm t}}

\newcommand{\betat}{{\beta}_{\rm t}}

\newcommand{\betai}{{\beta}_{\rm i}}

\newcommand{\deltai}{{\delta}_{\rm i}}

\newcommand{\rhoi}{{\rho}_{\rm i}}
\newcommand{\refer}{{\color{orange}REFERENCE}}

% Bibliography and bibfile
\def\aj{AJ}%
          % Astronomical Journal
\def\actaa{Acta Astron.}%
          % Acta Astronomica
\def\araa{ARA\&A}%
          % Annual Review of Astron and Astrophys
\def\apj{ApJ}%
          % Astrophysical Journal
\def\apjl{ApJ}%
          % Astrophysical Journal, Letters
\def\apjs{ApJS}%
          % Astrophysical Journal, Supplement
\def\ao{Appl.~Opt.}%
          % Applied Optics
\def\apss{Ap\&SS}%
          % Astrophysics and Space Science
\def\aap{A\&A}%
          % Astronomy and Astrophysics
\def\aapr{A\&A~Rev.}%
          % Astronomy and Astrophysics Reviews
\def\aaps{A\&AS}%
          % Astronomy and Astrophysics, Supplement
\def\azh{AZh}%
          % Astronomicheskii Zhurnal
\def\baas{BAAS}%
          % Bulletin of the AAS
\def\bac{Bull. astr. Inst. Czechosl.}%
          % Bulletin of the Astronomical Institutes of Czechoslovakia 
\def\caa{Chinese Astron. Astrophys.}%
          % Chinese Astronomy and Astrophysics
\def\cjaa{Chinese J. Astron. Astrophys.}%
          % Chinese Journal of Astronomy and Astrophysics
\def\icarus{Icarus}%
          % Icarus
\def\jcap{J. Cosmology Astropart. Phys.}%
          % Journal of Cosmology and Astroparticle Physics
\def\jrasc{JRASC}%
          % Journal of the RAS of Canada
\def\mnras{MNRAS}%
          % Monthly Notices of the RAS
\def\memras{MmRAS}%
          % Memoirs of the RAS
\def\na{New A}%
          % New Astronomy
\def\nar{New A Rev.}%
          % New Astronomy Review
\def\pasa{PASA}%
          % Publications of the Astron. Soc. of Australia
\def\pra{Phys.~Rev.~A}%
          % Physical Review A: General Physics
\def\prb{Phys.~Rev.~B}%
          % Physical Review B: Solid State
\def\prc{Phys.~Rev.~C}%
          % Physical Review C
\def\prd{Phys.~Rev.~D}%
          % Physical Review D
\def\pre{Phys.~Rev.~E}%
          % Physical Review E
\def\prl{Phys.~Rev.~Lett.}%
          % Physical Review Letters
\def\pasp{PASP}%
          % Publications of the ASP
\def\pasj{PASJ}%
          % Publications of the ASJ
\def\qjras{QJRAS}%
          % Quarterly Journal of the RAS
\def\rmxaa{Rev. Mexicana Astron. Astrofis.}%
          % Revista Mexicana de Astronomia y Astrofisica
\def\skytel{S\&T}%
          % Sky and Telescope
\def\solphys{Sol.~Phys.}%
          % Solar Physics
\def\sovast{Soviet~Ast.}%
          % Soviet Astronomy
\def\ssr{Space~Sci.~Rev.}%
          % Space Science Reviews
\def\zap{ZAp}%
          % Zeitschrift fuer Astrophysik
\def\nat{Nature}%
          % Nature
\def\iaucirc{IAU~Circ.}%
          % IAU Cirulars
\def\aplett{Astrophys.~Lett.}%
          % Astrophysics Letters
\def\apspr{Astrophys.~Space~Phys.~Res.}%
          % Astrophysics Space Physics Research
\def\bain{Bull.~Astron.~Inst.~Netherlands}%
          % Bulletin Astronomical Institute of the Netherlands
\def\fcp{Fund.~Cosmic~Phys.}%
          % Fundamental Cosmic Physics
\def\gca{Geochim.~Cosmochim.~Acta}%
          % Geochimica Cosmochimica Acta
\def\grl{Geophys.~Res.~Lett.}%
          % Geophysics Research Letters
\def\jcp{J.~Chem.~Phys.}%
          % Journal of Chemical Physics
\def\jgr{J.~Geophys.~Res.}%
          % Journal of Geophysics Research
\def\jqsrt{J.~Quant.~Spec.~Radiat.~Transf.}%
          % Journal of Quantitiative Spectroscopy and Radiative Trasfer
\def\memsai{Mem.~Soc.~Astron.~Italiana}%
          % Mem. Societa Astronomica Italiana
\def\nphysa{Nucl.~Phys.~A}%
          % Nuclear Physics A
\def\physrep{Phys.~Rep.}%
          % Physics Reports
\def\physscr{Phys.~Scr}%
          % Physica Scripta
\def\planss{Planet.~Space~Sci.}%
          % Planetary Space Science
\def\procspie{Proc.~SPIE}%
          % Proceedings of the SPIE
\let\astap=\aap
\let\apjlett=\apjl
\let\apjsupp=\apjs
\let\applopt=\ao

\newcommand{\figpath}{PFIGS/}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\title{SiO Molecular Jets around young stars - A numerical perspective}
\author[B. Vaidya, Tom Douglas, Paola Caselli, Tom Hartquist]{B. Vaidya$^{1}$\thanks{E-mail:
B.Vaidya@leeds.ac.uk (BV)}, Tom Douglas$^{1}$, Paola Caselli$^{1}$, Tom Hartquist$^{1}$\\
$^{1}$School of Physics and Astronomy, University of Leeds, Leeds LS2
9JT\\
}

\date\today

%\date{Accepted yyyy m dd. Received yyyy m dd; in original yyyy m dd}
\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{yyyy}

\maketitle

\label{firstpage}

\begin{abstract}
  % context heading (optional)
  {}  
  % aims heading (mandatory)
   {}
  % methods heading (mandatory)
   {}
  % results heading (mandatory)
   {}
  % conclusions heading (optional), leave it empty if necessary 
\end{abstract}

   \begin{keywords}
    MHD -- methods:numerical -- ISM: jets and outflows
   \end{keywords}

%
%

%===============================================================

\section{Introduction}
Jets are one of the first manifestations of star
formation in dense molecular cores. They are ubiquitous in both
massive and low star forming regions. These supersonic flows
perpendicular to the underlying accertion disk plays a vital role in
removing excess angular momentum and thereby aiding in the accretion. 
For low mass stars, they are believed to be launched by magneto-centrifugally forces and
further collimated by magnetic hoop stress
\citep[][]{Blandford:1982p892, Konigl:2000p607}. However, in case of
high mass stars, radiative forces also contribute to flow dynamics
during the later evolutionary stages \cite{Vaidya:2011p8992}. 
Typically, these jets are few parsec size long and can be divided into
three length-scale domains viz. source and disk scales (1-10$^{2}\,$AU),
envelope scales ($10^{2} - 10^{5}$\,AU) and parent cloud scales
($10^{5} - 10^{6}$\,AU) \refer. Among them, the envelope scales are
the ones where rich chemical evolution occurs as the jet interacts
with the molecular medium. In this region, the jet propagates into a
relatively static medium inducing shocks that are interesting from the
physical and chemical point of view. In addition to shocks, molecular
material from the surrounding is entrained and accelerated to high
velocities giving rise to molecular outflows.
%

Bipolar molecular outflows from low and high mass stars
have been studied in details over the past 
decade \citep[see reviews by][]{Bachiller:1996p4692,Arce:2007p798,
Tafalla:2011p14051}. Advancement in millimeter interferometers have
allowed to observe these outflows with high spatial resolution of
few arc seconds. A large number of studies related to young
outflows are done using standard outflow and shock tracer
like CO and SiO respectively. In addition to these molecular tracers,
shocks from these outflows are detected in molecular hydrogen using
infra-red telescopes. Based on these observational studies, various
empirical properties for these outflows have been discovered. For
example, the CO outflows from single dish studies are seen to be
highly collimated for both low and high mass stars
\citep[for e.g.,][]{Gueth:1999p4683, Beuther:2002p3574} they are
referred to as {\em{molecular jets}}. These outflows also exhibit a
mass-velocity relation with a power law form $ dM(v)/dv \propto
v^{-\gamma}$, where values of $\gamma$ range from 1 to 3 \cite{Downes:2003p9946}. Episodic
knots belived to be caused by variable accretion events are a common property
of young molecular outflows (for e.g. in L1577:
\citealt{Gueth:1998p14058}, in HH300: \citealt{Arce:2001p14064}). These knots show their signatures
as {\em{wedges}} in the position-velocity (PV) diagrams \cite{Arce:2001p14065}. Also, most
commonly observed in these outflows are signatures of rotation and
precision (for e.g., in DG Tau: \citealt{Bacciotti:2002p2084}).
%

Even with myriad of empirical evidences, the exact nature of SiO and
CO outflow is not clear. For a complete understanding it is imperative
to compliment these observations with theoretical models. Many models
based on hydro-simulations and steady state shock calculations
were proposed to explain the observational signatures of molecular
outflows. Among them the two main models are that of wide-angled wind
driven \cite{Shu:1991p14071} and jet driven outflow
\cite{Canto:1991p14123}. 
The most popular among them is the jet driven model as wind
driven molecular outflows not only fail to match observed PV digarams \cite{Cabrit:1992p14098}
but also tend to sweep large quantitity of material at the extermities
of the lobes \cite{Masson:1992p14101}. While the jet driven models could sucessfully
derive the global outflow shapes and mass velocity relations of CO
outflows \citep{Raga:1993p9948, Masson:1993p9661} . There have also been some attempts to combine these two
models into one \cite{Shang:2006p14268} to explain the global observational features. 
However, most of these dynamical models do not
account for shock chemsitry. Instead, shock chemistry is studied
independently without dynamics using plane parallel shock
models \citep{Schilke:1997p14140,Flower:2003p11236} . The magneto-hydrodynamic (MHD) calculations by \cite{Glassgold:1991p13703} suggested that molecules
like SiO and CO could as form within the jet. Similar conclusions of
molecules surving in steady state disk winds have also been shown
\cite{Panoglou:2012p10039}. Very limited simulations have modelled the outflow
dynamics including molecular chemistry but in absence of
magnetic fields \citep{Raga:1995p12965, Smith:2003p9985}.

%
In the present work, our goal is to take a step further in modelling of
jet driven molecular outflows. The present model aims to consistently derive observed emission
properties of molecular outflows, specifically various SiO line
tranistions, by combining axisymmetric MHD simulations of
radiative jet propogation. In particular, we evolve the jet dynamical
quantities in conjunction with different non-equilibrium cooling
prescriptions of varying complexities. The most complex is that of
molecular cooling along with H$_{2}$ chemistry. This prescription
allows us to track the formation and destruction of 
HI, HII and H2 along with the flow dynamics. 
The final state of the jet obtained for each cooling model is then
post-processed using a non-LTE radiative line transfer code
to obtain emission maps, spectra and PV digram. These emission maps
are further processed using ALMA-CASA pipeline to obtain synthetic
images of molecular outflows.
%

In the next three section we describe our numerical setup, cooling
presecriptions and radiative transfer line code respectively. In
Sect. 5, we will present results from the parameter survey and the
discussions along with predicted ALMA maps will be presented in
Section 6 and 7, followed by conclusions.

\section{Numerical Setup}
\subsection{Numerical code and Equations}
For our study, we carry out axisymmetric numerical ideal MHD simulations using the PLUTO code \citep{Mignone:2007p644} which is based on a conservative scheme of Godunov type.
We have modified the original code to incorporate molecular cooling
from self-consistent evolution of hydrogen chemistry (see Sect.~\ref{sec:chem}).

In general, the MHD code considers the following set of equations.
The conservation of the mass and the momentum,
%
\begin{equation}\label{masscons}
\frac{\partial \rho}{\partial t} + (\vec{v} \cdot \nabla)\rho  +
\rho \nabla \cdot \vec{v} = 0
\end{equation}
%
\begin{equation}\label{momcons}
\rho(\frac{\partial \vec{v}}{\partial t} +
(\vec{v} \cdot \nabla) \vec{v}) =
- \nabla P + \frac{1}{4\pi} (\nabla \times \vec{B}) \times \vec{B}
\end{equation}
%
where $\rho$ is gas density, $\vec{v}$ the velocity vector, $P$ the gas pressure,
and $\vec{B}$ the magnetic field vector with the poloidal and toroidal
components - $\vec{B}_{\rm p}, {B}_{\phi}$. Note that the forces due
to gravity are neglected for this problem as the domain of interest is
far away from the central object (i.e., star).  
%

The cooling function $\Lambda$ which dependents on temperature $T$, mass density $\rho$ and
chemical abundances {\bf{X}}, appears in the
energy equation as a source term,

\begin{equation}\label{encons1}
\frac{\partial}{\partial t}(\rho E)
+ \nabla \cdot\left[ \rho E \vec{v} + (P + \frac{B^2}{8\pi})\vec{v}\right]  
- \vec{B}(\vec{v}\cdot\vec{B}) = \Lambda(\rho,T,{\bf{X}}),
\end{equation}
%
%\begin{equation}\label{encons1}
%\frac{\partial}{\partial t}(\rho {\rm E}) + \nabla.\left[ (\rho
%{\rm E})\textbf{v} + (\textbf{P} + \frac{B^2}{8
%\pi})\textbf{v}\right] - \textbf{B}(\textbf{v.B}) = \rho\left[
%-\nabla\Phi + (\textbf{F}^{\rm rad})\right].\textbf{v}
%\end{equation}
%
where the total energy density of the flow $E$ comprises contributions from 
the internal energy $\epsilon$, the mechanical energy and the magnetic energy,
%
\begin{equation}\label{encons2}
 E = \epsilon + \frac{v^2}{2} + \frac{B^2}{8 \pi \rho}.
\end{equation}
The gas pressure in the flow is related to the density assuming an equation 
of state with the adiabatic index $\gamma$,
%
\begin{equation}\label{EOS}
P = (\gamma - 1) \rho \epsilon.
\end{equation}

The evolution of chemical abundances for each species is solved via,
%
\begin{equation}\label{chemevol}
\frac{\partial \rho{\bf{X}}_{i}}{\partial t} + \nabla \cdot (\rho
{\bf{X}}_{i} \vec{v})  = \rho {\bf{S}}_{i},
\end{equation}
where ${\bf{S}}_{i}$ represents the net creation or destruction of a
given species through chemical reactions (see Sect.~\ref{sec:chem}).

The evolution of the magnetic field is governed by induction equation,
%
\begin{equation}\label{induction}
\frac{\partial \vec{B}}{\partial t} = \nabla \times \left(\vec{v}\times \vec{B}\right).
\end{equation}
%
In addition to the above set of equations the code obeys the condition of divergence-free 
magnetic fields, $\nabla \cdot \vec{B} = 0$, which is numerically achieved by construction 
since using the Powell's eight wave formulation.
\subsection{Initial Condition}
We model the propagation of jet as it interacts with the molecular
cloud core much further from the central object, i.e., $>$ 1000\,AU. 
Further away from the central source, the downward pull of gravity plays a
negligible role and the dynamics of jet is primarily governed by magnetic fields.
The total magnetic field in jet is dominated by the toroidal
component. This is because the poloidal field decays as $z^{-2}$ as
compared to $z^{-1}$ for toroidal field to maintain the force balance ($z$ is the vertical distance from source). 
The hoop stress due to pinch force from toroidal magnetic fields maintains the
a highly collimated beam like structure for the jet.
%

The ambient medium with which the jet interacts primarily
represents the molecular cloud core. The numerical domain is axi-symmetric
and in (r,z,$\phi$) cylindrical co-ordinates. Its extent in radial
direction is 20 $R_{\rm j}$ and 100
$R_{\rm j}$ along the vertical axis, $R_{\rm j}$ being the radius of
the jet. The domain is resolved by an uniform grid with 200 cells in
radial and 1000 cells in vertical direction. For simplicity, we choose this medium
to be unmagnetized and non-turbulent. The density in the ambient
medium varies with vertical height $z$ as, $\rho_{\rm amb}\sim (\rho_0/z^{2})$
consistent with observations \citep{Caselli:2011p13935}. The value of
$\rho_0$ depends upon the density contrast, $\eta$, between the jet and
the ambient medium. The number density in the jet is kept fixed such
that the density in ambient medium lies within a range of $10^{4}-10^{5}
\rm cm^{-3}$. The pressure in the ambient molecular medium is
set so to maintain a constant temperature of 50\,K. 
%

The jet enters into the medium through a nozzle of radius $R_{\rm jet}$
from the lower boundary (z = 0). The jet density is fixed to
be $10^{5} \rm cm^{-3}$ and it has a radius of $2.5\times10^{15} \rm cm \sim
167\,AU$. The jet is injected into the domain with a typical 
velocity of $v_{\rm jet,0}$ = 100 km/s as is the case for most low mass stellar jets
specially the low velocity component. The constant jet velocity is
superimposed with periodic pulsation of the form,
\begin{equation}
v_{\rm jet} = v_{\rm jet,0}(1.0 + A\,sin(2.0\pi t/T_{\rm p}))
\end{equation}
where the amplitude A is 0.25 and time period $T_{\rm p} = 70$
years. The pressure at the surface of the jet is $10^{-10} \rm
dyne\,\,cm^{-2}$ corresponding to a temperature of $T_{\rm jet} \sim
4\times10^{3}\,K$. Inside the jet beam, a radial variation of thermal
and magnetic pressure is adopted to maintain a magneto-static
equilibrium. We adopt the
same radial profiles used by \cite{Stone:2000p2650} for all our runs. 
Based on these profiles, the toroidal magnetic field is assumed to be
zero at the axis and achieves a maximum at some radius, $r_{\rm m}$
inside the jet. The maximum value, $B_{\phi,\rm m}$, depends on the
plasma $\beta$ which is a parameter used for our study along with the
density contrast $\eta$.
%

In order to consistently model the SiO emission arising from shocks as
this jet interacts with the medium, we evolve the dynamics along with
chemistry and cooling prescribtions. They are described in details in
the next section.

\section{Chemistry and Cooling}
\label{sec:chem}
\subsection{Power law cooling}
\subsection{Atomic Cooling}
\subsection{Tabulated Cooling}

\subsection{Molecular Cooling}
\label{ssec:molcool}
The evolution of molecular, atomic and ionized hydrogen is governed by
equations listed in Table~\ref{tab:chemeq}. In this cooling mode,
these equations are evolved at each times using temperature dependent
rates mentioned in the table along with their source. The code tracks
the formation and destruction of three quantities viz., X(HI), X($H_{2}$)
and X(HII) with a constraint that sum of all three should be unity.
Further, these abundances are used to update the cooling
function $\Lambda(n,T,{\bf{X}})$ to consistently derive the
temperature for next advection step.

\begin{figure*}
 \includegraphics[width=1\columnwidth]{\figpath/H2ReactionRates.png}
 \includegraphics[width=1\columnwidth]{\figpath/H2CoolingFuncs.png}
 \caption{Variation of $H_2$ chemistry reaction rates, $k_{i}$ and cooling
   function $\Lambda(n,T,{\bf{X}})$ with temperature for the initial
   state (see Sect.~\ref{ssec:molcool})}
\label{fig:tempvar}
\end{figure*}

\begin{table*}
\begin{minipage}{\textwidth}
\caption{Summary of the chemistry reaction set. T is the temperature
  in Kelvin, $T_{\rm eV}$ is the temperature in electron-volts, $T_{5}$
  = $T/1\times10^{5}$  and 
$T_{2}$  = T/100}
\label{tab:chemeq}
\begin{tabular}{l l l l}
\hline
No. & Reaction & Rate Coefficient ($\rm {cm}^{3} s^{-1}$) &
Reference~\footnote{REFERENCES -- (1) \cite{Cen:1992p13616} [Eq. 26a];
  (2) \cite{Woodall:2007p13623} [UMIST Database] (3)
  \cite{Galli:1998p13066} [Eq. H17]; (4) \cite{Abel:1997p12836}
  [Tab. 3 Eq. 13]; (5) \cite{Hollenbach:1979p12707} [Eq. 3.8]}\\
\hline
1. & H + e$^{-}$ $\rightarrow$ H$^{+}$ + 2e$^{-}$ & $k_1$ = $5.85
\times 10^{-11}$ $T^{0.5}$ \rm{exp}(-157,809.1/T)/(1.0 + $T_{5}^{0.5}$) & 1\\
2. & H$^{+}$ + e$^{-}$ $\rightarrow$ H + h$\nu$ & $k_2$ =
$3.5\times10^{-12} (T/300.0)^{-0.8}$ & 2\\
3. & H$_{2}$ + e$^{-}$ $\rightarrow$ 2H + e$^{-}$ & $k_3$ =
$4.4\times10^{-10} T^{0.35} \rm{exp}(-102,000.0/T)$ & 3\\
4. & H$_{2}$ + H $\rightarrow$ 3H & $k_4$ = $1.067\times10^{-10}
T_{\rm eV}^{2.012}(\rm{exp}(4.463/T_{\rm eV})^{-1}((1. + 0.2472 T_{\rm eV})^{3.512})^{-1} $& 4\\
5. &H$_{2}$ + H$_{2}$ $\rightarrow$ H$_{2}$ + 2H & $k_5$ = $1.0\times 10^{-8} \rm{exp}(-84,100/T)$ & 2\\
6. & H + H $\overset{\rm dust}\longrightarrow$ H$_{2}$ & $k_6 =
3.0\times10^{-17}\sqrt{T_{2}}(1.0 + 0.4\sqrt{T_{2} + 0.15} + 0.2T_{2} + 0.8T_{2}^{2})$ & 5 \\
\hline
\end{tabular}
\end{minipage}
\end{table*}

\section{Radiative Transfer}
Radiative transfer modeling used for post processing.

\subsection{The radiative transfer code} \label{subsec:radiative_transfer_code}
The radiative transfer program used is LIME (LIne Modeling Engine; \citealt{Brinch:2010p13078}), which  calculates line intensities based on a weighted sample of randomly chosen points in a continuous 3D model. The method of selecting these points is given in section \ref{subsec:gridding}. At each of these points, the density of the main collision partner (equivalent amount of H$_2$, given by n(H$_2$)+ 0.5 n(H)), gas and dust temperatures, velocity, molecular abundances and unresolved turbulent velocity are specified. These points are then smoothed by Lloyd's algorithm \citep{Lloyd1982} in order to minimise the variation in distance between points whilst keeping the same underlying distribution. These points are then connected by Delaunay triangulation and it is between the points connected by this method that photon are allowed to propagate (fig.~\ref{grid}). The level populations of the selected molecules are calculated at each of these points from collisional and radiative (de)excitation and the local radiation field is calculated. This is repeated 20 times with the populations of each level converging towards a single value. This number of iterations is sufficient for the signal to noise ratio of the level populations (as defined in \citealt{Brinch:2010p13078}) to exceed 1000 for 99\% of the points, ensuring that the simulation has converged on a stable level population. After 20 iterations the model is ray-traced in order to produce synthetic brightness maps. The average of ten separate runs was taken to minimise the artefacts in the output images, resulting from the grid construction (Fig.~\ref{points}).

\begin{figure}
 \includegraphics[width=84mm]{\figpath/eta2Beta10.png}
 \caption{A plot of the points selected by the gridding process and the paths down which photons can propagate for points in the r,z plane. The points are color coded by the density distribution (in m$^{-3}$, as used in LIME) and are more concentrated in the high density knots.}
\label{grid}
\end{figure}

\subsection{Grid construction} \label{subsec:gridding}
In order to construct the grid, candidate points are randomly selected from the volume to be simulated. These candidates then have their equivalent H$_2$ number density, and the number density of SiO, compared against those of a reference point in order to decide if the candidate point is to be used in the grid or not. Candidate grid points are selected at random in a cylindrical coordinates that is linearly spaced in z and $\theta$ and logarithmically spaced in r. For each point to be selected, a random number $\alpha$ is drawn from the semi-open set [0,$\,$1) as a threshold. After selection of random coordinates, the H$_2$ density and SiO density at the candidate point (n and m, respectively) are compared against the densities of a reference point in the un-perturbed ambient medium multiplied by $\frac{4\eta}{5}$ (n$_0$ and m$_0$). If $\alpha<\left( \frac{n}{n_0} \right)^{0.3}$ or $\alpha< \left( \frac{m}{m_0} \right)^{0.3}$ then the point is selected for use. Otherwise another r, $\theta$, z co-ordinate is selected and it becomes the candidate point. In addition to this method of selection, 5\% of the points are linearly distributed in x, y and z with no bias with regards to density or abundance. This provides a minimum level of sampling for the large low density regions in the outer parts of the simulated volume. See fig. \ref{points} for an example of the points distribution in r, z. The function comparing the candidate point to the reference point and the candidate point distribution were selected empirically to sample all scales while ensuring that the majority of points are located in the inner disc where the density is higher.  \smallskip

\begin{figure}
 \includegraphics[width=63mm]{\figpath/eta2Beta10_histo.png}
 \caption{A 2D histogram of the point distribution throughout the model. The disc and envelope can be seen as two separate entities which have to be sampled using different point distributions.}
 \label{points}
\end{figure}

\subsection{SiO abundance}
Need refs for this!!\smallskip
The amount of SiO is determined by the local velocity and temperature. The fractional abundance is given by the equation:\smallskip

$log(X)=-2.48\times10^{-8}v^5 + 5.50\times10^{-6}v^4 - 4.28\times10^{-4}v^3 + 1.24\times10^{-2}v^2 + 2.52\times10^{-2}v - 1.20\times10^{1}$\smallskip

where v is in kilometres per second. In addition to this if the temperature at the point is greater than 92,000K (the temperature of the Si-O bond disassociation energy) the abundance is reduced to 10$^{-30}$.

\begin{figure}
 \includegraphics[width=84mm]{\figpath/abun.png}
 \caption{The polynomial used to calculate an abundance of SiO as a function of local velocity.}
 \label{abun}
\end{figure}

\section{Parameter Survey}
\begin{table*}
\centering
\caption{Summary from parameter runs}
\begin{tabular}{l | l | l | l | l | l }
\hline
Run & Cooling Mode & $\eta$ & $\beta$ & Peak Intensity [K] & FWHM at Bow
Shock \\
\hline\hline
adi1010 & Nil (Adiabatic) & 10 & 10 &&\\
pow1010 & Power law & 10 & 10 &&\\
atm1010 & Atomic & 10 & 10 &&\\
atm101 & Atomic & 10 & 1 &&\\
atm210 & Atomic & 2 & 10 &&\\
tab1010 & Tabulated & 10 & 10 &&\\
tab210 & Tabulated & 2 & 10&&\\
mol1010 & Molecular & 10 & 10&&\\
mol510 & Molecular & 5 & 10&&\\
\hline
\end{tabular}
\label{tab:result1}

\end{table*}

\section{Results}
\subsection{Comparison of cooling prescription}
figure nos. 2 -- Similuated images, emission maps for eta = 10.
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/jetplt1_1010.png}
 \caption{Jet Volume Density for different cooling modes with
   $\eta$ = 10 and $\beta$ = 10.}
\label{grid}
\end{figure*}

\begin{figure}
 \includegraphics[width=84mm]{\figpath/J2-1_mom0maps_cooling.png}
 \caption{A plot of the integrated SiO J2-1 emission from 5 models, 
each using a different method to calculate cooling and all with $\eta$=10 $\beta$=10.}
\label{cooling2-1}
\end{figure}

\subsection{Molecular cooling and H$_2$ Chemistry}
figure nos. 2 --> Explain molecular cooling in more details. With the
mechanism of formation and destruction of HI, H2 and HII using
figures. 
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/molcjet_310.png}
 \caption{Fraction of hydrogen species in the run with molecular
   cooling having $\eta$ = 3 and $\beta$ = 10.}
\label{grid}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/molcjet_310_molform.png}
 \caption{Dependence of hydrogen fractions on the temperature at two
   points in the flow, viz. the interface of the knot with molecular
   medium and at the bow shock}
\label{grid}
\end{figure*}

\subsection{Effects of changing $\eta$ and $\beta$}
figure nos. 2 -- Simulated images describing lengths, collimation and
emission for lower eta more stronger. 
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/jetplt2_diffetabeta.png}
 \caption{Jet Volume Density for atomic cooling with $\eta$ = 10 but $\beta$ = 1 and
   10 and that with tabulated cooling with $\beta$ = 10 but $\eta$ = 2
   and 10.}
\label{grid}
\end{figure*}

\subsection{SiO Abundance profile}
figure nos. 2 -- Comparison of three profiles for SiO Abundance,
Contour maps for two transitions for top hat profile.

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/contourmapsSiOTopHat.png}
 \caption{Contours of SiO 2-$>$1 and 8-$>$7 emission obtained using the
   SiO Top Hat profile.}
\label{grid}
\end{figure}

\subsection{Spectra and PV diagrams}
figure nos. 1 -- one transition 3 different angles, Comparison of
spectra and line ratios

Table 1 -- Parameter runs with changing $\beta$ (magnetic fields) and
$\eta$ (density contrast) for cooling \\
Results should have Peak Intensity after convolution., FWHM at tip of
bow shock.\\
Our reference run -- $\beta$ = 10.0, $\eta$ = 2.0, with cooling and
velocity dependent SiO Abundance. \\
Figure 3 -- Output from MHD simulations (Reference run).\\
Figure 4 -- Output from Rad Transfer (Reference run). Integrated
Intensity Maps for different transitions for three different cooling prescribtions. (TODO : TOM)\\
Figure 5 -- Spectra for the reference runs all transitions. (TODO :
TOM)\\
Figure 6 -- PV diagrams (TODO : TOM) for one transition in two directions.
Figure 7 -- Angle of Inclination dependence

\section{Discussion}
Effects of varying $\eta$\\
Lower $\eta$ gives better emission.

Importance of cooling prescriptions\\
Bulk of emission comes from cooling instabilities.

Effects of varing abundance profile ... Crucial ingredient in
modeling molecular bullets \\

Spectra and PV Digrams\\
Line - withs $\sim$ 15-25 km\,s$^{-1}$. Wedge shaped PV diagrams
a signature feature of jet driven molecular outflows. Counterparts in observations.

Comparison with Observed results \\
Different line transition trace different components of SiO
emission. The low level transitions trace the region where the jet
interacts with molecular medium. The clumpy nature in its emission is
from the turbulent nature of its interaction. The higher transitions
trace a much more collimated region along with the terminal shock. 
As seen by Chandler \& Richer 2001 and Hirano 2006.
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/linejets.png}
 \caption{SiO emission from different line transitions for the run with molecular
   cooling having $\eta$ = 3 and $\beta$ = 10. }
\label{grid}
\end{figure*}

Advantage of non-LTE modeling over LVG Cabrit 2007, Lee 2008. 
Line ratios $\sim$ 1 can be explained well. \\

EHV component ?? \\

Limitations\\

\section{ALMA view}
ALMA view of the reference run and stress of applying our synthetic techniques
to study the molecular outflows in more details.\\
Figure 6. for ALMA -- \\

\section{Conclusion}
We are the best in modelling SiO outflows.

%\section*{Acknowledgments}

\bibliographystyle{mn2e}
\bibliography{bibfile}
\label{lastpage}
\end{document}