### 2013-12-22 TALK/TifrTalk.tex

\documentclass[8pt,xcolor=dvipsnames]{beamer}
\usepackage[latin1]{inputenc}

\usetheme[height=7mm]{Berlin}
\usecolortheme[named=Brown]{structure}

\usepackage{time}       % date and time
\usepackage{graphicx}
\usepackage[T1]{fontenc}    % european characters
%\usepackage{courier}
\usepackage{animate}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{amssymb,amsmath}  % use mathematical symbols
\usepackage{bookman}                  % use palatino as the default font
\setbeamercovered{transparent}
\newcommand{\figpath}{/home/phybva/SiOJets_New/PAPER/NEWFIGS/}

\title[YSO Outflows]{Outflows from Young Stellar Objects}
\subtitle{MHD, Radiation \& Chemistry}
\author[Bhargav Vaidya]{\textcolor{blue}{Bhargav Vaidya}\inst{1}}
\institute[Uni. Leeds]{\inst{1}School of Physics and Astronomy, University of Leeds, Leeds.}
\date[2014]{Visitor Seminar TIFR 2014}
%\titlegraphic{\vskip30pt \hskip215pt\includegraphics[width=0.3\textwidth]{\figpath/Leeds-logo.jpg}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\frametitle{Outline}
\tableofcontents
\end{frame}

\section{Introduction}
\begin{frame}{Star formation : What do we know}

\end{frame}

\begin{frame}{Star formation : Feedback}
\end{frame}

\begin{frame}{Present challenges}
\end{frame}

\section{Motivation}
\begin{frame}{Outflow Evolutionary Picture}
\end{frame}

\begin{frame}{Chemistry in outflows}
\end{frame}

\begin{frame}{Molecular bullets and EHV emission}
\end{frame}

\section{Methods : Numerical Simulations}
\begin{frame}{Numerical code}
\end{frame}

\begin{frame}{Launching and Propagation}
\end{frame}

\begin{frame}{Radiation force}
\end{frame}

\begin{frame}{Chemistry and Cooling}
\end{frame}

\section{Outflow Dynamics : Launching}
\begin{frame}{MHD Acceleration}
\end{frame}

\begin{frame}{Collimation and Radiation}
\end{frame}

\begin{frame}{Force Parameters and its impact}
\end{frame}

\begin{frame}{Resistive effects}
\end{frame}

\begin{frame}{A case of Orion Source I}
\end{frame}

\section{Outflow dynamics : Propagation}
\begin{frame}{Molecular Interplay}
\end{frame}

\begin{frame}{Spectra and PV digrams}
\end{frame}

\begin{frame}{Multi-Line survey : Emission}
\end{frame}

\begin{frame}{Multi-line survey : Line Ratios}
\end{frame}

\begin{frame}{Focussing on a single knot}
\end{frame}

\begin{frame}{Rotation or Wiggles?}
\end{frame}

\section{Summary}
\begin{frame}{Conclusions}
\end{frame}

\end{document}